```{r}
library(tidyverse)
library(forcats)
library(GGally)
library(cowplot)
```

```{r}
eeg = read.csv("EEG.machinelearing_data_BRMH.csv")
mood = eeg %>% filter(main.disorder == "Healthy control" | main.disorder == "Mood disorder")
ab = eeg[, 1:122]
eeg %>% head
nrow(eeg)
nrow(mood)
```

```{r}
eeg %>% 
  mutate(label = ifelse(main.disorder == "Healthy control", 
                        1, ifelse(main.disorder == "Mood disorder", 
                                  2, 3))) %>% 
  ggplot(aes(y = fct_infreq(main.disorder), fill = label)) +
  geom_bar() +
  labs(y = "Main Disorder", title = "Count of numbers of main disorder in the dataset") +
  theme_bw() + 
  theme(legend.position = "none")
```


```{r,fig.height=6,fig.width=6}
eeg %>% 
  ggplot(aes(y = fct_infreq(specific.disorder))) +
  geom_bar() +
  facet_wrap("main.disorder", scale = "free_y", nrow = 4) +
  labs(y = "Specific Disorder", title = "Count of numbers of main disorder in the dataset") +
  theme_bw()
```

```{r,fig.height=2,fig.width=6}
sex.p = eeg %>% ggplot(aes(x = sex)) + geom_bar() + theme_bw()
iq.p = eeg %>% ggplot(aes(x = IQ)) + geom_histogram() + theme_bw()
edu.p = eeg %>% ggplot(aes(x = education)) + geom_histogram() + theme_bw()
age.p = eeg %>% ggplot(aes(x = age)) + geom_histogram() + theme_bw()

plot_grid(sex.p, iq.p, edu.p, age.p, nrow = 1)
```

```{r}
mood %>% select(main.disorder, age, IQ, education, sex) %>% summary
```

```{r}
mood %>% select(main.disorder, age, IQ, education, sex) %>% 
  ggpairs
```

Now categorizing EEG signal:

reference: https://www.kaggle.com/code/lazygene/visualising-pre-processed-eeg-data 

columns to be in format: band.channel\
reformat from XX.X.band.x.channel to band.channel

AB: absolute power per band per channel (eg. AB.A.delta.a.FP1)
COH: coherence

We only look at AB here. 


```{r}
ab = eeg[, 1:122]
colnames(ab)[9:122] = str_split(colnames(ab)[9:122], "\\.") %>% 
  lapply(function(x){ paste0(x[3], "_", x[5])
  }) %>% unlist
ab.long = ab %>% pivot_longer(9:122) %>% 
  separate(name, c("band", "chanel"), "_") %>% 
  filter(main.disorder == "Healthy control" | main.disorder == "Mood disorder")
ab.long %>% head
```

in each band: correlation between chanels?\
in each chanel: correlation between bands?

band (6): "delta" "theta" "alpha" "beta" "highbeta" "gamma" \
chanel (19):  "FP1" "FP2" "F7"  "F3"  "Fz"  "F4"  "F8"  "T3"  "C3"  "Cz" "C4"  "T4"  "T5"  "P3"  "Pz"  "P4"  "T6"  "O1"  "O2"  

```{r}
d = ab.long %>% filter(band == "delta") %>% 
  pivot_wider(names_from = chanel, values_from = value) %>% 
  select(10:28) %>% ggcorr()
t = ab.long %>% filter(band == "theta") %>% 
  pivot_wider(names_from = chanel, values_from = value) %>% 
  select(10:28) %>% ggcorr()
a = ab.long %>% filter(band == "alpha") %>% 
  pivot_wider(names_from = chanel, values_from = value) %>% 
  select(10:28) %>% ggcorr()
b = ab.long %>% filter(band == "beta") %>% 
  pivot_wider(names_from = chanel, values_from = value) %>% 
  select(10:28) %>% ggcorr()
h = ab.long %>% filter(band == "highbeta") %>% 
  pivot_wider(names_from = chanel, values_from = value) %>% 
  select(10:28) %>% ggcorr()
g = ab.long %>% filter(band == "gamma") %>% 
  pivot_wider(names_from = chanel, values_from = value) %>% 
  select(10:28) %>% ggcorr()

plot_grid(d, t, a, b, h, g, nrow = 2,
          labels = c("delta", "theta", "alpha", "beta", 
                     "highbeta", "gamma"))
```

chanel (19):  "FP1" "FP2" "F7"  "F3"  "Fz"  "F4"  "F8"  "T3"  "C3"  "Cz" "C4"  "T4"  "T5"  "P3"  "Pz"  "P4"  "T6"  "O1"  "O2"  

```{r}
ab.long %>% filter(chanel == "F8") %>% 
  pivot_wider(names_from = band, values_from = value) %>% 
  select(10:15) %>% ggcorr()
```

